Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-11-15 grove-paper.tex
add code view screenshots
add my orcid
auto-format whitespace
VSCode insists, so I'll wrap it in a pull request, in case the changes
get in the way.
make code in prose match the diagrams
also fix a few small grammar errors and add some whitespace to latex
source

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureSimple
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{SimpleTimes}{Root}%
\newvertex{SimpleTimes}{*}%
\newedge{SimpleX}{L}%
\newvertex{SimpleX}{x}%
\newcommand{\figureSimple}{
\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:A simple tree with a hole-a}
TODO
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:A simple tree with a hole-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style, color=black] -> {
  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Code that Includes a Hole.
\autoref{fig:A simple tree with a hole-a}:~User Interface for the Code.
\autoref{fig:A simple tree with a hole-b}:~Graph Representation of the Code.}
\label{fig:A simple tree with a hole}
\end{figure}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureMultiChild
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MultiChildAlice}{R}
\newvertex{MultiChildAlice}{x}
\newedge{MultiChildBob}{R}
\newvertex{MultiChildBob}{y}
\newcommand{\figureMultiChild}{
\begin{figure*}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested-c}}}};
%%%%
\path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
\end{tikzpicture}
}
\label{fig:Multi-child-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {
    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
\end{tikzpicture}
}
\label{fig:Multi-child-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
\path [draw,->,merge step] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
\end{tikzpicture}
}
\label{fig:Multi-child-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'anchor=-15] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
   {},
   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{}
\label{fig:Multi-child-d}
TODO: show UI
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Multi-Child Conflicts.}
\label{fig:Multi-child}
\end{figure*}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureMultiParent
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MultiParentAlice}{R}
\newedge{MultiParentBob}{R}
\newcommand{\figureMultiParent}{
\begin{figure*}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-child-c}}}};
%%%%
\path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-child-c-x) to[star] (-2em, 0cm |- fig-Multi-parent-a);
\end{tikzpicture}
}
\label{fig:Multi-parent-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    {}
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-parent-a) to[star] (-2em, 0cm |- fig-Multi-parent-b);
\end{tikzpicture}
}
\label{fig:Multi-parent-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    "\vNestedAlice" [> "\eMultiParentAlice"', >alice edge]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Multi-parent-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-parent-c);
\end{tikzpicture}
}
\label{fig:Multi-parent-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'],
   "\vNestedAlice" [> "\eMultiParentBob", >bob edge]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
\path [draw,->,merge step] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
\end{tikzpicture}
}
\label{fig:Multi-parent-d}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge edge]
   },
   {}
  }
 },
};
\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
\end{tikzpicture}
}
\label{fig:Multi-parent-e}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    "\vNestedAlice" [> "\eMultiParentAlice"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{}
\label{fig:Multi-parent-f}
TODO: UI
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Multi-Parent Conflicts.}
\label{fig:Multi-parent}
\end{figure*}
}

@@

TODO: UI

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureDisconnection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{DisconnectionAlice}{R}
\newedge{DisconnectionBob}{L}
\newcommand{\figureDisconnection}{
\begin{figure*}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-User Cycles-d}}}};
%%%%
\path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-User Cycles-d-x) to[star] (-2em, 0cm |- fig-Disconnection-a);
\end{tikzpicture}
}
\label{fig:Disconnection-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'],
   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Disconnection-a) to[star] (-2em, 0cm |- fig-Disconnection-b);
\end{tikzpicture}
}
\label{fig:Disconnection-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    "\vMultiCycleTimes" [> "\eDisconnectionAlice", >alice edge]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Disconnection-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Disconnection-c);
\end{tikzpicture}
}
\label{fig:Disconnection-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
    {}
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
\path [draw,->,merge step] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
\end{tikzpicture}
}
\label{fig:Disconnection-d}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\linewidth}
\centering
\caption{}
\label{fig:Disconnection-e}
TODO: UI
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Disconnection.}
\label{fig:Disconnection}
\end{figure*}
}

@@

\author{Eric Griffis}
\orcid{nnnn-nnnn-nnnn-nnnn}

@@

Finally, holes are represented by the absence of a child.
For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \vSimpleTimes{}
corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.

@@

To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole-b}
and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
This action corresponds to creating \eConstructionY{} from \vSimpleX{} at the \texttt{R} position to a newly
created vertex, \vConstructionY{}, containing the variable reference~\texttt{y}.
The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.

@@

Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
This is modeled by deleting \eConstructionY{} as shown in \autoref{fig:Change-a}. 
Notice that \vConstructionY{} continues to exist, and if it had any children, those children would remain connected to it.
In our system, once a vertex is created it is never deleted.
This allows further manipulations of those vertices by other users as shown later in this section.
Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.

Once an edge with a particular identifier is deleted, it cannot be recreated.
Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.

@@

@@

To start, Alice and Bob both have the graph in \autoref{fig:Move-b}, which
represents the code~\texttt{\hole{} + (\hole{} + y)}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureSimple
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{SimpleTimes}{Root}%
\newvertex{SimpleTimes}{*}%
\newedge{SimpleX}{L}%
\newvertex{SimpleX}{x}%
\newcommand{\figureSimple}{
\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:A simple tree with a hole-a}
\includegraphics{code-views/simple.png}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:A simple tree with a hole-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style, color=black] -> {
  "\vSimpleTimes" [> "\eSimpleTimes", >color=black, color=black] -> {
   "\vSimpleX" [> "\eSimpleX"', >color=black, color=black],
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Code that Includes a Hole.
\autoref{fig:A simple tree with a hole-a}:~User Interface for the Code.
\autoref{fig:A simple tree with a hole-b}:~Graph Representation of the Code.}
\label{fig:A simple tree with a hole}
\end{figure}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureMultiChild
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MultiChildAlice}{R}
\newvertex{MultiChildAlice}{x}
\newedge{MultiChildBob}{R}
\newvertex{MultiChildBob}{y}
\newcommand{\figureMultiChild}{
\begin{figure*}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Editing Nested-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Editing Nested-c}}}};
%%%%
\path (0cm,0cm) node (fig-Multi-child conflict-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Editing Nested-c-x) to (-2em, 0cm |- fig-Multi-child conflict-a);
\end{tikzpicture}
}
\label{fig:Multi-child-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {
    "\vMultiChildAlice" [> "\eMultiChildAlice", >alice edge, alice node]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-child conflict-b) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Editing Nested-c-x) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-b);
\end{tikzpicture}
}
\label{fig:Multi-child-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   "\vMultiChildBob" [> "\eMultiChildBob", >bob edge, bob node]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-child conflict-c) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Multi-child conflict-a) to[out=15,in=165] (-2em, 0cm |- fig-Multi-child conflict-c);
\path [draw,->,merge step] (fig-Multi-child conflict-b) to (-2em, 0cm |- fig-Multi-child conflict-c);
\end{tikzpicture}
}
\label{fig:Multi-child-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'anchor=-15] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   "\vMultiChildAlice" [> "\eMultiChildAlice"anchor=153, >merge edge, merge node],
   {},
   "\vMultiChildBob" [> "\eMultiChildBob"anchor=195, >merge edge, merge node]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.21\linewidth}
\centering
\caption{}
\label{fig:Multi-child-d}
\includegraphics{code-views/multi-child-conflict.png}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Multi-Child Conflicts.}
\label{fig:Multi-child}
\end{figure*}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureMultiParent
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{MultiParentAlice}{R}
\newedge{MultiParentBob}{R}
\newcommand{\figureMultiParent}{
\begin{figure*}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Multi-child-c-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-child-c}}}};
%%%%
\path (0cm,0cm) node (fig-Multi-parent-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-child-c-x) to[star] (-2em, 0cm |- fig-Multi-parent-a);
\end{tikzpicture}
}
\label{fig:Multi-parent-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    {}
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-b) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-parent-a) to[star] (-2em, 0cm |- fig-Multi-parent-b);
\end{tikzpicture}
}
\label{fig:Multi-parent-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    "\vNestedAlice" [> "\eMultiParentAlice"', >alice edge]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-c) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Multi-parent-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Multi-parent-c);
\end{tikzpicture}
}
\label{fig:Multi-parent-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'],
   "\vNestedAlice" [> "\eMultiParentBob", >bob edge]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-d) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Multi-parent-b) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
\path [draw,->,merge step] (fig-Multi-parent-c) to[out=15,in=165] (-2em, 0cm |- fig-Multi-parent-d);
\end{tikzpicture}
}
\label{fig:Multi-parent-d}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  a/{\vWrapPlus} [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    b/"\vNestedAlice" [> "\eMultiParentAlice"', >merge edge]
   },
   {}
  }
 },
};
\path [draw,->,merge edge] (a) to ["\eMultiParentBob",out=-60,in=60] (b);
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Multi-parent-e) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-parent-d) to (-2em, 0cm |- fig-Multi-parent-e);
\end{tikzpicture}
}
\label{fig:Multi-parent-e}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  v8/"\vWrapPlus" [> "\eWrapPlus"] -> {
   v2/"\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    "\vNestedAlice" [> "\eMultiParentAlice"']
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.14\linewidth}
\centering
\caption{}
\label{fig:Multi-parent-f}
\includegraphics{code-views/multi-parent-conflict.png}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Multi-Parent Conflicts.}
\label{fig:Multi-parent}
\end{figure*}
}

@@

\includegraphics{code-views/cycle.png}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \figureDisconnection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newedge{DisconnectionAlice}{R}
\newedge{DisconnectionBob}{L}
\newcommand{\figureDisconnection}{
\begin{figure*}
%\hfill
\hskip0.12\linewidth
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (-7.5em,0cm) node (fig-Multi-User Cycles-d-x) [gray,anchor=base west] {{\textbf{\small Fig.~\ref*{fig:Multi-User Cycles-d}}}};
%%%%
\path (0cm,0cm) node (fig-Disconnection-a) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Multi-User Cycles-d-x) to[star] (-2em, 0cm |- fig-Disconnection-a);
\end{tikzpicture}
}
\label{fig:Disconnection-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'],
   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Disconnection-b) [anchor=base] {\strut};
\path [draw,->,alice step] (fig-Disconnection-a) to[star] (-2em, 0cm |- fig-Disconnection-b);
\end{tikzpicture}
}
\label{fig:Disconnection-b}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    {},
    "\vMultiCycleTimes" [> "\eDisconnectionAlice", >alice edge]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Disconnection-c) [anchor=base] {\strut};
\path [draw,->,bob step] (fig-Disconnection-a) to[out=15,in=165,star] (-2em, 0cm |- fig-Disconnection-c);
\end{tikzpicture}
}
\label{fig:Disconnection-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vMultiCycleTimes" [> "\eMultiCycleAliceTimes"] -> {
    "\vSimpleTimes" [> "\eDisconnectionBob"', >bob edge],
    {}
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\textwidth}
\centering
\caption{
\begin{tikzpicture}[remember picture, overlay]
\path (0cm,0cm) node (fig-Disconnection-d) [anchor=base] {\strut};
\path [draw,->,merge step] (fig-Disconnection-b) to[out=15,in=165] (-2em, 0cm |- fig-Disconnection-d);
\path [draw,->,merge step] (fig-Disconnection-c) to (-2em, 0cm |- fig-Disconnection-d);
\end{tikzpicture}
}
\label{fig:Disconnection-d}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (-0.5cm,-2cm) graph[graph style] { plus/"\vSimpleTimes" };
\path ( 0.5cm,-2cm) graph[graph style] { times/"\vMultiCycleTimes" };
\path[draw,->,merge edge] (times) to ["\eMultiCycleBobTimes"'anchor=north,out=-135,in=-45] (plus);
\path[draw,->,merge edge] (plus) to ["\eMultiCycleBobPlus"anchor=south,out=45,in=135] (times);
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.17\linewidth}
\centering
\caption{}
\label{fig:Disconnection-e}
\includegraphics{code-views/disconnected-reachable.png}
\includegraphics{code-views/disconnected-unreachable.png}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Example of Disconnection.}
\label{fig:Disconnection}
\end{figure*}
}

@@

\author{Eric Griffis}
\orcid{0000-0003-1693-6172}

@@

Finally, holes are represented by the absence of a child.
For example, in \autoref{fig:A simple tree with a hole-b} the absence of a right edge coming from \vSimpleTimes{}
corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.

@@

To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole-b}
and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
This action corresponds to creating \eConstructionY{} from \vSimpleTimes{} at the \texttt{R} position to a newly
created vertex, \vConstructionY{}, containing the variable reference~\texttt{y}.
The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.

@@

Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
This is modeled by deleting \eSimpleX{} as shown in \autoref{fig:Change-a}.
Notice that \vSimpleX{} continues to exist, and if it had any children, those children would remain connected to it.
In our system, once a vertex is created it is never deleted.
This allows further manipulations of those vertices by other users as shown later in this section.
Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.

Once an edge with a particular identifier is deleted, it cannot be recreated.

Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \vSimpleTimes{} and \vSimpleX{}.

@@

\newcommand{\eWrapTimesWrapC}{\eWrapTimes}

@@

To start, Alice and Bob both have the graph in \autoref{fig:Move-b}, which
represents the code~\texttt{\hole{} + (\hole{} * y)}.
